---
  title: "Project"
---
  
  I am using GSS data to explore income by sex. It is my hypothesis that men earn more than women and I expect this to be true. This is an interesting problem because men earn more than women in most parts of the world but there are many factors such as education and health that could play a role in this. Income is a relevant issue for nearly anyone, so I expect this hypothesis to pique the interest of many readers. 
  
  According to GSS data, more women than men earn in every income bracket except $4,000 to $4,999. This means that for the data collected by GSS, women largely earn more than men. The variables used to determine this were "sex" and "income." "Sex" refers to the sex of the GSS respondent where 1 is a man and 2 is a woman.  "Income" refers to total family income bracket where 1 refers to an income under $1,000, and ranges to 12 which refers to an income over $25,000. 
    The mean income for men ("sex" 1) was about $19,000 with a standard deviation of 2.4. The mean income for women ("sex" 2) was over $25,000 with a standard deviation of 2.3. 
    
```{r}
sd(gss2$income[gss2$sex == 1], na.rm = TRUE) 
```
and
```{r}
sd(gss2$income[gss2$sex == 2], na.rm = TRUE)
```
were used to determine the standard deviation for men and women respectively while mean was determined using 
```{r}
mean(gss2$income[gss2$sex == 1], na.rm = TRUE)
```
and
```{r}
mean(gss2$income[gss2$sex == 2], na.rm = TRUE)
```
for men and women respectively.

  Using a Chi-squared test with 11 degrees of freedom, we find the Chi-squared value to be 15.255 and the p-value to be 0.1711. This tells us that if there were no correlation between sex and income, the data would exist 17.11% of the time. In other words, there is a 17.11% (relatively low) probability that deviation from the expected results is due to chance only. The R code used to determine this is as follows: 
  
  ```{r}
chisq.test(gss$sex, gss$income)
```

  It does need to be noted that more women than men were surveyed in this study, which could inflate the responses of women to appear stronger than those of men. Therefore multiple Chi-squared tests were run. The code below was used to determine that there is indeed a positive correlation between education and income as well as between health and income.
  
  
```{r}
chisq.test(gss$educ, gss$income)
```
and 
```{r}
chisq.test(gss$health, gss$income)
```

